\pkg{plotROC}: A Better Tool for Plotting ROC Curves

### 2015-10-14 Paper/sachs2015plotROC.Rmd
motivation

There are several \proglang{R} packages related to ROC curve estimation that contain dedicated plotting functions. The \pkg{ROCR} package [@rocr] plots the FPF _versus_ TPF, as usual, and then takes the interesting approach of encoding the cutoff values as a separate color scale along the ROC curve itself. A legend for the color scale is placed along the vertical axis on the right of the plotting region. Due to the popularity of \pkg{ROCR}, my package supports the performance objects as returned by \pkg{ROCR}. The \pkg{pROC} package [@pROC] is mainly focused on estimating confidence intervals and regions for restricted ranges of the ROC curve. The plotting methods therein use the base \proglang{R} plotting functions to create nice displays of the curves along with shaded confidence regions. My \pkg{plotROC} package uses the \pkg{ggplot2} [@ggplot2] plotting library to create clear, informative ROC plots, with interactive features for use on the web, and sensible defaults for use in print.

## Motivation

@@

# C
plot(TPF ~ FPF, data = R1, type = 's', lwd  = 2, xlab = "False positive fraction", ylab = "True positive fraction", main = "C")
abline(v = seq(0, 1, by = .2), lwd = 1, col = "grey80")
abline(h = seq(0, 1, by = .2), lwd = 1, col = "grey80")
abline(0, 1, col = "grey80")
lines(TPF ~ FPF, data = R1, type = 's', lwd = 2)
text(.2, .8, "Test A")

@@

>>>>

There are several \proglang{R} packages related to ROC curve estimation that contain dedicated plotting functions. The \pkg{ROCR} package [@rocr] plots the FPF _versus_ TPF, as usual, and then takes the interesting approach of encoding the cutoff values as a separate color scale along the ROC curve itself. A legend for the color scale is placed along the vertical axis on the right of the plotting region. Due to the popularity of \pkg{ROCR}, my package provides a convenience function to plot ROC curves from the performance objects as returned by \pkg{ROCR}. The \pkg{pROC} package [@pROC] provides an option for plotting cutoff labels (`print.thres = TREU`) and is mainly focused on estimating confidence intervals and regions for restricted ranges of the ROC curve. The plotting methods therein use the base \proglang{R} plotting functions to create nice displays of the curves along with shaded confidence regions. My \pkg{plotROC} package uses the \pkg{ggplot2} [@ggplot2] plotting library to create clear, informative ROC plots, with interactive features for use on the web, and sensible defaults for use in print.

## Motivation and Literature Review

@@

# C
plot(TPF ~ FPF, data = R1, type = 's', lwd  = 2, xlab = "False positive fraction", ylab = "True positive fraction", main = "C", xlim = c(-.05, 1))
abline(v = seq(0, 1, by = .2), lwd = 1, col = "grey80")
abline(h = seq(0, 1, by = .2), lwd = 1, col = "grey80")
abline(0, 1, col = "grey80")
lines(TPF ~ FPF, data = R1, type = 's', lwd = 2)
text(.35, .6, "Test A")
dexc <- seq(1, length(R1$FPF), length.out = 7)[-c(1, 7)]
text(R1$FPF[dexc] - .075, R1$TPF[dexc] + .05, round(R1$c[dexc], 1))
lines(TPF ~ FPF, data = R1[dexc, ], type = 'p', pch = 20)

@@

The vast majority of the figures that were reviewed looked more like those in panels A and B than those in C and D. While plots like this do technically display the trade-offs between false- and true-positives, they minimize the amount of useful information that can be displayed in the space needed to plot an ROC curve. The plots created by \pkg{plotROC} attempt to increase the amount of information displayed in ROC charts that would otherwise be mostly white space. This is useful not only for print media, where space is limited, but also during data analysis. The analyst can quickly and easily view information that would otherwise be obscured by standard plotting software. The interactive features take this one step further, enhancing the plots with high density and easily accessible supplementary information.
